---
title: "Case Study 2 Deliverable TAJAR"
author: "Tori Wheelis, Andy Ho, An Nguyen, Jodi Pafford, Rajat Chandra"
date: "8/4/2018"
output:
  html_document:
    keep_md: yes
---
# Predicting And reducing Attrition Using Data Science

```{r, echo=TRUE}
# In order to import XLSX, we install the package "readxl" into our environment.  After loading library "readxl" we're able to use read_excel function to import the sheet "HR-employee-attrition Data" into a dataframe. 
library(ggplot2)
library(readxl)
library(randomForest)
library(stringr)
library(plyr)
library(dplyr)
library(tidyr)
library(xlsx)
library(reshape2)
library(party)
library(gmodels)
library(vcd)
library(caret)
library(e1071)
library(tm)
library(SnowballC)
library(wordcloud)
library(RColorBrewer)
sessionInfo()
attritionDF <- read_excel(path = "Attrition_Data_Clean.xlsx", sheet = "HR-employee-attrition Data")
```

## Introduction

Attrition, or voluntary employee turnover, comes as a high cost to companies.  The cost of interviewing, training, productivity, and negative impacts on morale are just a few. The key to success for companies, then, is to attract and retain talent. To help facilitate talent management and reduce attrition, thereby reducing real cost to companies, DDSAnalytics's in-house data science team, TAJAR, has compiled some important characteristics associated with increased risk of employee turnover. Our findings will allow companies to identify possible causes of employee turnover, create targeted approaches, and improve their talent management along with their company culture and profit margins.

DDSAnalytics is an analytics company that specializes in talent management solutions for Fortune 1000 companies. Talent management, broadly defined as the iterative process of developing and retaining employees, can include workforce planning, employee training programs, identifying high-potential employees and reducing/preventing voluntary employee turnover (attrition). DDSAnalytics has joined forces with TAJAR, their new in-house data science team, to apply data analysis to a dataset from IBM in order to help improve IBM's attrition rate, and more broadly, the attrition rate of Fortune 1000 companies. This report gives a picture of the employees in the dataset, the top factors for attrition, and suggests targeted ways for companies to mitigate their attrition rates.
 

## Methodology

This data, obtained from IBM's corporate office, allows a snapshot in to a large, thriving corporate company whose profit margins and general company culture could benefit from reducing attrition rates. After the data was cleaned and corrected for any formatting errors (since a few of the measures were self-reports from the employees), TAJAR reviewed the descriptive statistics. Then, TAJAR built a random forest model and edited it to identify the variables most strongly associated with attrition. After the initial findings, TAJAR continued to dig deeper and understand the "why" behind the numbers.

TAJAR used, in addition to some basic descriptive statistics, a random forest model to identify the most salient factors contributing to employee attrition (xxRAJAT**-- add more detail here. Expand on what you put in the slides if you started there first).

All of the code used to explore the data, including exploratory code not attached to this report, more details on the research methodology, and the data, as well a copy of this report, can be found online at https://github.com/R-Chandna/MSDS6306_CaseStudy_2.git.

```{r, echo=TRUE}
#1a	Tori: The client wants this to be reproducible and know exactly what you did.  There needs to be an informative Readme, complete with several sections, as referenced in Live Session.  Give contact information, session Info, and the objective of the repo at least.  
#1a: The README was updated to look pretty and have the basic introduction to everything. 
#1b	An: You have a large data set, and it needs its own Codebook, formatted in an approachable way.  Make sure you describe peculiarities of the data by variable and what needs transforming.  However, do not make it too long either.
#1b: The codebook was added as a separate file and to the README
#1c	Rajat: Create a file structure that is accessible and transparent.  Document it in the root directory, ideally in the Readme.
#1c: A clearly understandable directory was created and documented in the README **RAJAT**
```


## Dataset Overview, Demographics

The dataset used came without any missing values, and required only a little cleaning, mostly renaming the variables to be more readable and workable, and changing the format of a few variables so the data imported cleanly.

```{r, echo=TRUE}
library(stringr)

# 2A - a	Read the csv into R and take a look at the data set.  Output how many rows and columns the data.frame is.
# There are 1470 rows (observations) and 35 columns (variables)
dimDF <- dim(attritionDF)
names(dimDF) <- c("Rows", "Columns")
dimDF

# 2B - b	The column names are either too much or not enough.  Change the column names so that they do not have spaces, underscores, slashes, and the like. All column names should be under 12 characters. Make sure you’re updating your codebook with information on the tidied data set as well.
# Results of column names and its string length count.  
## Need to find a more elegant function to filter the list and subset only values greater than 12.  For now, this will suffice.
lettercount <- str_length(names(attritionDF))
countDF <- data.frame(cbind(names(attritionDF), lettercount))
names(countDF) <- c("Variable", "LetterCount")
countDF

# Replacing Original Variable names that have > 12 characters with new names under the Working column in Attrition_Col_Name_New.csv.

original <- names(attritionDF)
write.csv(original, "Attrition_Col_Name_Old.csv")
head(names(attritionDF))
colNamesDF <- read.csv("Attrition_Col_Name_New.csv")
working <- colNamesDF$Working
names(attritionDF) <- working
head(names(attritionDF))


# 2C - Due to Qualtrics, the following columns will have its integer value replaced by the definitions provided in the "Data Definitions" sheet.  
## Will need to seperate the integer and characters into two seperate columns.
datadefDF <- read_excel ("CaseStudy2-data.xlsx", "Data Definitions")
datadefDF

# 2D - d	Make sure your columns are the proper data types (i.e., numeric, character, etc.).  If they are incorrect, convert them. 
# Verifying the columns class type are numeric and character.
as.matrix(lapply(attritionDF, class))
```

The dataset included information gathered on 1,470 employees who voluntarily filled out an online survey and had 35 variables that included whether or not they left the company (attrition), demographics such as age, gender, and marital status, as well as employee information such as job role, stock options, and compensation, and also included some other metrics, such as relationship, environment, and job satisfaction, etc. There is a codebook included with this report that details all 35 variables and their names as used in our code.

```{r, echo=TRUE}
table(attritionDF$Gender)
sort(unique(attritionDF$JobRole))
table(attritionDF$Department)
```
Of the employees, 882 were male and 588 were female. The job roles listed were Healthcare Representative, Human Resources, Laboratory Technician, Manager, Manufacturing Director, Research Director, Research Scientist, Sales Executive, and Sales Represantative spread across Human Resources with 63 employees, Research & Development with 961 employees, and Sales with 446 employees. Of Human Resource's 63 employees, 11 were managers. Of R&D's 961, 54 were managers, and of Sales' 446, 37 were managers.

```{r, echo=TRUE}
#3c. Give the frequencies (in table format or similar) for Gender, Education, and Occupation.  They can be separate tables, if that's your choice.
#Needs some refining
gendertable <- table(attritionDF$Gender)
gendertable

educationtable <- table(attritionDF$Education)
educationtable

occupationtable <- table(attritionDF$JobRole)
occupationtable

#3d. Give the counts (again, table) of management positions.
management <- count(attritionDF, Department, JobRole)
management
filter(management, JobRole=="Manager")

#3b - b	Please provide (in table format or similar), descriptive statistics on at least 7 variables (age, Income, etc.).  Create a simple histogram for two of them.  Comment on the shape of the distribution in your markdown.
# Create a table of summary statistics of Variables with numerical values
SummaryStat <- matrix(ncol = 6, nrow = 0)
SummaryStat <- rbind(SummaryStat, summary(attritionDF$MonthlyInco))
SummaryStat <- rbind(SummaryStat, summary(attritionDF$Age))
SummaryStat <- rbind(SummaryStat, summary(attritionDF$DistFromHome))
SummaryStat <- rbind(SummaryStat, summary(attritionDF$PctSalaryInc))
SummaryStat <- rbind(SummaryStat, summary(attritionDF$TotWorkExp))
SummaryStat <- rbind(SummaryStat, summary(attritionDF$YrsAtCompany))
SummaryStat <- rbind(SummaryStat, summary(attritionDF$NumPriorComp))
SummaryStat <- as.data.frame(SummaryStat)
row.names(SummaryStat) <- c("MonthlyInco", "Age", "DistFromHome", "PctSalarInc", "TotWorkExp", "YrsAtCompany", "NumPriorComp")
SummaryStat
```

```{r, echo=TRUE}
table(attritionDF$Attrition)

#Department
ggplot(attritionDF, aes(x = Department, fill = Attrition)) + geom_bar(position = "fill") + labs(title = "Attrition per Department", x = "", y = "", color = "") + scale_y_continuous(labels = scales::percent) + scale_fill_manual(values=c(brewer.pal(11, "Spectral")[3], brewer.pal(11, "Spectral")[10]))
tempDF <- attritionDF[,c("Attrition","Department")]
table(melt(tempDF))

#Age
ggplot(attritionDF, aes(x = Age, fill = Attrition)) + geom_histogram(position = "fill") + labs(title = "Attrition per Age Groups", x = "Age (years)", y = "", color = "") + scale_y_continuous(labels = scales::percent) + scale_fill_manual(values=c(brewer.pal(11, "Spectral")[3], brewer.pal(11, "Spectral")[10]))

#Income
ggplot(attritionDF, aes(x = MonthlyInco, fill = Attrition)) + geom_histogram(position = "fill") + labs(title = "Attrition per Monthly Income", x = "USD", y = "", color = "") + scale_y_continuous(labels = scales::percent) + scale_fill_manual(values=c(brewer.pal(11, "Spectral")[3], brewer.pal(11, "Spectral")[10]))

```
Out of the 1470 employees, 237 voluntarily left the company. The department with the most employees who left was the the Sales department with ~20.6% attrition rate followed by Human Resources with ~19.0% and R&D with ~13.8%. Employees between the ages of 18-21, who did not have stock options, who were less satisfied with their jobs, with lower incomes, who worked in sales, and who had the lowest incomes also had higher attrition rates. However, there seemed to be no difference in attrition according to gender.

```{r, echo=TRUE}
#Create table with summary of variables with categorial variables, using first two factors most likely to attribute to attrition
JobSatTable <- as.data.frame(table(attritionDF$JobSatIndex))
JobSatTable <- data.frame(JobSatTable[,-1], row.names = JobSatTable[,1])
colnames(JobSatTable) <- c("Freq")

#Barplot of Job Satisfaction Level
ggplot(JobSatTable, aes(reorder(x=row.names(JobSatTable), -Freq), y=Freq)) + geom_bar(stat = "identity", aes(fill = row.names(JobSatTable))) + labs(title = "Employees' Job Satisfaction", y = "Number of Employees", x = "") + theme(axis.text.x = element_text(angle = 0, hjust = 1), legend.position="none") + scale_fill_brewer(palette="Spectral")

OverTimeTable <- as.data.frame(table(attritionDF$OverTime))
OverTimeTable <- data.frame(OverTimeTable[,-1], row.names = OverTimeTable[,1])
colnames(OverTimeTable) <- c("Freq")

#Barplot of Over Time
ggplot(OverTimeTable, aes(reorder(x=row.names(OverTimeTable), -Freq), y=Freq)) + geom_bar(stat = "identity", aes(fill = row.names(OverTimeTable))) + labs(title = "Over Time", y = "Number of Employees", x = "") + theme(axis.text.x = element_text(angle = 0, hjust = 1), legend.position="none") + scale_fill_brewer(palette="Spectral")

#Age
ggplot(attritionDF, aes(x = Age, fill = Attrition)) + geom_bar(position = "fill") + labs(title = "Age and Attrition", x = "Age", y = "", color = "")+ scale_y_continuous(labels = scales::percent)
```

When looking at Life Satisfaction, we looked at the participants who were no longer working at the company against the self-reporting Work Life Balance rating of "Bad", "Good", "Better", or "Best" and the Job Satisfaction Index of "Low", "Medium", "High", and "Very High".

```{r Satisfaction, echo=TRUE}
#Create a Life Satisfaction data frame with Attrition (Yes), Work Life Balance and Job Satsifaction.
Satisfaction <- subset(attritionDF, Attrition=="Yes", select=c(JobSatIndex, WorkLifeFit, Attrition))
head(Satisfaction)

#Create a visual to display.
ggplot(Satisfaction, aes(x = WorkLifeFit, fill = JobSatIndex)) + geom_bar(position = "fill") + labs(title = "Life Satisfaction - Attrition", x = "work-Life Balance with Job Satisfaction", y = "", color = "")+ scale_y_continuous(labels = scales::percent) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + scale_x_discrete(limits=c("Bad","Good","Better","Best"))


#Create a Life Satisfaction data frame with Attrition (No), Work Life Balance and Job Satsifaction.
SatisfactionN <- subset(attritionDF, Attrition=="No", select=c(JobSatIndex, WorkLifeFit, Attrition))
head(SatisfactionN)

#Create a visual to display.
ggplot(SatisfactionN, aes(x = WorkLifeFit, fill = JobSatIndex)) + geom_bar(position = "fill") + labs(title = "Life Satisfaction - Not Lost to Attrition", x = "Work-Life Balance with Job Satisfaction", y = "", color = "")+ scale_y_continuous(labels = scales::percent) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + scale_x_discrete(limits=c("Bad","Good","Better","Best"))
```

While exploring the data, these trends related to job role (and that may or may not directly relate to attrition) surfaced:

```{r, echo=TRUE}

#Job Satisfaction vs Job Role in %
attritionDF$JobSatIndex <- factor(attritionDF$JobSatIndex, levels = c("Low", "Medium", "High", "Very High"))
ggplot(attritionDF, aes(JobRole, fill = JobSatIndex)) + geom_bar(position = "fill")  + labs(title = "Job Satisfaction per Job Role" , y = "", x = "") + theme(axis.text.x = element_text(angle = 55, hjust = 1)) + scale_fill_brewer(palette="Spectral") + scale_y_continuous(labels = scales::percent)

#Over Time vs Job Satisfaction in %
ggplot(attritionDF, aes(JobSatIndex, fill = OverTime)) + geom_bar(position = "fill")  + labs(title = "Over Time and Job Satisfaction" , y = "", x = "") + theme(axis.text.x = element_text(angle = 0, hjust = 1)) + scale_fill_brewer(palette="Spectral") + scale_y_continuous(labels = scales::percent)
attritionDF$JobSatIndex <- as.character(attritionDF$JobSatIndex)
```

For privacy reasons, IBM and TAJAR with DDSAnalytics did not want to run any analyses on minors. There were, however, no minors in the dataset. The age of the participating employees ranged from 18-60.


```{r, echo=TRUE}
#3a -	Remove all observations where the participant is under age 18.  No further analysis of underage individuals is permitted by your client.  Remove any other age outliers as you see fit, but be sure to tell what you’re doing and why.
# Age between 18 and 60, no children under 18 and no obvious age outliers.
summary(attritionDF$Age)
attritionDF$Over18[attritionDF$Over18 == 'N']
```

## Deeper Analysis, Evaluation, Results

Once TAJAR completed the random forest, it was found that the top 7 factors contributing to attrition were: 1) How satisfied an employee was with their hob, 2) Whether an employee reported working overtime, 3) How satisfied an employee was with their general environment, 4) What level their job was considered, on a scale from 1-5, 5) an employee's hourly rate, 6) An employee's work/life balance score, 7) and Whether or not the employee was married (which was closely correlated with their age.)

```{r Attrition_DataSet_Read}
# Create a copy of original DataFrame
formattedAttrDF <- data.frame(attritionDF)
str(formattedAttrDF)
```

```{r Correct_Format_Of_Col_In_DataFrame}
as.data.frame(apply(formattedAttrDF, 2, function(x) length(unique(x))))
# Converting relevant Character Variables to Factors since these have levels. 
convertToFactors <- c("Attrition", "WorkTravel", "Department", "Education", "FieldOfStudy", "EnvrSatIndex", "Gender", "JobInvolv", "JobLevel", "JobRole", "JobSatIndex" , "MaritalState", "Over18", "OverTime", "PerfRating", "RelpSatIndex", "StockOptLvl", "WorkLifeFit")
for(columnName in convertToFactors){
  formattedAttrDF[,columnName] <- as.factor(formattedAttrDF[,columnName])
}
# Extra Debug--Remove
lapply(formattedAttrDF, function(x){if(is.factor(x)) levels(x)})
levelsInfoDf <- read.xlsx2(file = "CaseStudy2-data.xlsx",sheetName = "Data Definitions", startRow = 3, header = F)
str(levelsInfoDf)
levelsInfoDf[ , c("levelNo", "levelName")] <- colsplit(levelsInfoDf$X2, " ", c("levelNo", "levelName"))
levelsInfoDf$levelName <- gsub('[[:punct:]]+','',levelsInfoDf$levelName)
levelsInfoDf

# Give Meaningful Category Names to Levels of factors that have levels as "1", "2" etc 
# This will be done based upon the data definations present in the case study sheet
levels(formattedAttrDF$Education) <- levelsInfoDf[1:5, "levelName"]
levels(formattedAttrDF$EnvrSatIndex) <- levelsInfoDf[7:10, "levelName"]
levels(formattedAttrDF$JobInvolv) <- levelsInfoDf[12:15, "levelName"]
levels(formattedAttrDF$JobSatIndex) <- levelsInfoDf[17:20, "levelName"]
levels(formattedAttrDF$PerfRating) <- levelsInfoDf[22:25, "levelName"]
levels(formattedAttrDF$RelpSatIndex) <- levelsInfoDf[27:30, "levelName"]
levels(formattedAttrDF$WorkLifeFit) <- levelsInfoDf[32:35, "levelName"]
# Assigning Levels to variables Job Level and Stock Options as these are not mentioned in data defination sheet
levels(formattedAttrDF$JobLevel) <- c("Entry", "Experienced", "First-Level Management", "Middle-Level Management", "Top-Level Management")
levels(formattedAttrDF$WorkTravel) <- c("Non-Travel", "Travel_Rarely", "Travel_Frequently")
lapply(formattedAttrDF, function(x){if(is.factor(x)) levels(x)})

all(complete.cases(formattedAttrDF))
```

```{r doingALittleEDA}
## Create a vector of all categorical variables
categoricalVarVec <- c("WorkTravel", "Department", "Education", "FieldOfStudy", "EnvrSatIndex", "Gender", "JobInvolv", "JobLevel", "JobRole", "JobSatIndex" , "MaritalState", "OverTime", "PerfRating", "RelpSatIndex", "StockOptLvl", "WorkLifeFit")
for(categoricalVar in categoricalVarVec){
  CrossTable(formattedAttrDF[ ,categoricalVar], formattedAttrDF$Attrition, chisq = T)
  mosaicplot(CrossTable(formattedAttrDF[ ,categoricalVar], formattedAttrDF$Attrition)$t, main = paste("Attrition Vs", categoricalVar, sep = " "), xlab = categoricalVar, ylab = "Attrition", color = T)
}
## Do the same for Numeric Variables
numericVarVec <- c("Age", "DailyRate", "DistFromHome", "HourlyRate", "MonthlyInco", "MonthlyRate", "NumPriorComp", "PctSalaryInc" , "TotWorkExp", "TrainedLstYr", "YrsAtCompany", "YrsInRole", "LastPromoted", "YrsWithMgr")
# Recode the Attrition Var
formattedAttrDF$RecodedAttrVar <- ifelse(formattedAttrDF$Attrition == "Yes", 1, 0)
for(numericVar in numericVarVec){
  plot(x = formattedAttrDF[ ,numericVar], y = formattedAttrDF$RecodedAttrVar,main = paste("Attrition Vs", numericVar, sep = " "), xlab = numericVar, ylab = "Attrition")
}
# Drop the Recoded Attrition Column
formattedAttrDF <- subset(formattedAttrDF, select = -RecodedAttrVar)
```

```{r Fitting_The_Model}
set.seed(99)
subformattedAttrDF <- subset(formattedAttrDF, select = -EmployeeID)
ModelRandomForest <- randomForest(Attrition ~ . , data=subformattedAttrDF, proximity=T)
ModelRandomForest
```

```{r Plotting_OOBError_Vs_No_Of_Trees}
oob.error.data <- data.frame(
  Trees=rep(1:nrow(ModelRandomForest$err.rate), times=3),
  Type=rep(c("OOB", "No", "Yes"), each=nrow(ModelRandomForest$err.rate)),
  Error=c(ModelRandomForest$err.rate[,"OOB"],
    ModelRandomForest$err.rate[,"No"],
    ModelRandomForest$err.rate[,"Yes"]))

ggplot(data=oob.error.data, aes(x=Trees, y=Error)) +
  geom_line(aes(color=Type))

# Adding More Trees to Forest and checking Out of Bag Error Rate
ModelRandomForestWith1000Trees <- randomForest(Attrition ~ . , data=subformattedAttrDF, ntree = 1000,  proximity=T)
ModelRandomForestWith1000Trees

oob.error.data <- data.frame(
  Trees=rep(1:nrow(ModelRandomForestWith1000Trees$err.rate), times=3),
  Type=rep(c("OOB", "No", "Yes"), each=nrow(ModelRandomForestWith1000Trees$err.rate)),
  Error=c(ModelRandomForestWith1000Trees$err.rate[,"OOB"],
    ModelRandomForestWith1000Trees$err.rate[,"No"],
    ModelRandomForestWith1000Trees$err.rate[,"Yes"]))

ggplot(data=oob.error.data, aes(x=Trees, y=Error)) +
  geom_line(aes(color=Type))
```

```{r Determining_The_Appropriate_Mtry_Value}
outOfBagErrorValues <- vector(length = (dim(subformattedAttrDF)[2] - 1))
for (randomVarCount in 1:length(outOfBagErrorValues)){
  testModel <- randomForest(Attrition ~ . , data=subformattedAttrDF, ntree = 500,  mtry = randomVarCount)
  outOfBagErrorValues[randomVarCount] <- testModel$err.rate[nrow(testModel$err.rate),1]
}
outOfBagErrorValues
plot(outOfBagErrorValues, type = "o", col= "Blue")
```

```{r Fitting_The_FinalModel_BasedOn_Mtry_And_nTree_Values}
# Appropriate ntree and mtry values for lowest out of bag error are as 500 and 11. Hence,
# fitting final model based on these
FinalFittedModel <- randomForest(Attrition ~ . , data=subformattedAttrDF, ntree = 500, mtry = 11)
FinalFittedModel
```

```{r Finding_And_Plotting_Top_Predictors}
importanceMatrix <- importance(FinalFittedModel)
importanceMatrix[order(-importanceMatrix[,1]), ]
varImpPlot(FinalFittedModel)
```

```{r Finding_And_Plotting_Top_Predictors_Conditional_RForest}
subformattedAttrDF <- subset(subformattedAttrDF, select = -Over18)
cf1 <- cforest(Attrition ~ . , data=subformattedAttrDF, control=cforest_unbiased(mtry=11,ntree=500))
#relativeImp_CondRandForest <- varimp(cf1, conditional=TRUE)
#relativeImp_CondRandForest <- relativeImp_CondRandForest[order(-relativeImp_CondRandForest[,1]), , drop = F]
#relativeImp_CondRandForest
# Importance based upon decrease in acuracy without corelation
relativeImp_CondRandForest <-  varimpAUC(cf1)
write.csv(relativeImp_CondRandForest, file = "Testing4.csv", row.names = T, append = F)
# Importance based upon decrease in acuracy with corelation
relativeImp_CondRandForest_1 <-  varimpAUC(cf1, conditional = T)
#relativeImp_CondRandForest_1 <- relativeImp_CondRandForest_1[order(-relativeImp_CondRandForest_1[,1]), , drop = F]
relativeImp_CondRandForest_1
write.csv(relativeImp_CondRandForest_1, file = "Testing3.csv", row.names = T, append = F)
# Calculate Model Accuracy
caret:::cforestStats(cf1)
```


```{r runLogisticRegressionModel}
names(subformattedAttrDF)
subformattedAttrDF <- subset(subformattedAttrDF, select = -c(StdHours, Headcount))
logisticModelFit <- glm(Attrition ~ ., data=subformattedAttrDF, family = "binomial")
summary(logisticModelFit)
relativeImp <- varImp(logisticModelFit, scale = F)
relativeImp <- relativeImp[order(-relativeImp[,1]), , drop = F]
relativeImp
```

**So both Conditional Random Forest and Logistic Regression Model have found Overtime, JobSatisfaction and EnvironmentSatisfaction as top three predictors.**

```{r creatTheWordCloud}
wordCloudDF <- read.xlsx(file = "WordCloudData.xlsx", header = F, 1)
wordcloud(words = wordCloudDF$X1, freq = wordCloudDF$X2, random.order=FALSE, rot.per=0, colors=brewer.pal(8, "Dark2"))
```



Since, during the demgraphics and exploratory phase of the data analysis, employees between the ages of 18-21 seemed to have a significant spike in attrition, and income seemed to have some relationship with attrition, TAJAR decided to further explore the relationship of these variables. It was found that age and income are correlated with each other. Age, though mostly normally distributed throughout the company, has a slightly larger amount of employees around the age of 36. The incomes of these employees, however are not normally distributed and a large majority of the employees make at or below 5,000 dollars a month, or 60,000 dollars a year or less.

```{r, echo=TRUE}
#An's code for Q4b
library(RColorBrewer)

#4b there seemed to be no relationship between age and MonthlyRate, DailyRate, or HourlyRate
##Monthly Rate
ggplot(data = attritionDF, aes(x = Age, y = MonthlyRate)) + geom_point(aes(colour = factor(Gender))) + geom_smooth(method = "lm", aes(group = Gender, colour = Gender)) + labs(title = "Montly Rate vs Age", x = "Age", y = "Monthly Rate", color = "Gender")
test <- lm(attritionDF$MonthlyRate ~ attritionDF$Age)
summary(test)
##Daily Rate
ggplot(data = attritionDF, aes(x = Age, y = DailyRate)) + geom_point(aes(colour = factor(Gender))) + geom_smooth(method = "lm", aes(group = Gender, colour = Gender)) + labs(title = "Daily Rate vs Age", x = "Age", y = "Daily Rate", color = "Gender")
test <- lm(attritionDF$DailyRate ~ attritionDF$Age)
summary(test)
##Hourly Rate
ggplot(data = attritionDF, aes(x = Age, y = HourlyRate)) + geom_point(aes(colour = factor(Gender))) + geom_smooth(method = "lm", aes(group = Gender, colour = Gender)) + labs(title = "Hourly Rate vs Age", x = "Age", y = "Hourly Rate", color = "Gender")
test <- lm(attritionDF$HourlyRate ~ attritionDF$Age)
summary(test)

#4b. MonthlyInco shows correlation with Age.  Tried linear regression on untransformed data and log, reciprocal, square root transformations
## untransformed
ggplot(data = attritionDF, aes(x = Age, y = MonthlyInco)) + geom_point(aes(colour = Gender)) + geom_smooth(method = 'lm', aes(group = Gender, colour = Gender)) + labs(title = "Montly Income vs Age", x = "Age", y = "Monthly Income", color = "Gender")
## log transformed
ggplot(data = attritionDF, aes(x = Age, y = log(MonthlyInco))) + geom_point(aes(colour = Gender)) + geom_smooth(method = 'lm', aes(group = Gender, colour = Gender)) + labs(title = "Montly Income vs Age", x = "Age", y = "log(Monthly Income)", color = "Gender", subtitle = "log transformed")
## reciprocal transformed
ggplot(data = attritionDF, aes(x = Age, y = -1/MonthlyInco)) + geom_point(aes(colour = Gender)) + geom_smooth(method = 'lm', aes(group = Gender, colour = Gender)) + labs(title = "Montly Income vs Age", x = "Age", y = "-1/(Monthly Income)", color = "Gender", subtitle = "negative reciprocal transformed")
## square root transformed
ggplot(data = attritionDF, aes(x = Age, y = sqrt(MonthlyInco))) + geom_point(aes(colour = Gender)) + geom_smooth(method = 'lm', aes(group = Gender, colour = Gender)) + labs(title = "Montly Income vs Age", x = "Age", y = "sqrt(Monthly Income)", color = "Gender", subtitle = "square root transformed")

#4b. Transformation did not yield better fit, fit test performed on most uncomplicated model, untransformed data
test <- lm(attritionDF$MonthlyInco ~ attritionDF$Age, subset = attritionDF$Gender == 'Male')
summary(test)
test <- lm(attritionDF$MonthlyInco ~ attritionDF$Age, subset = attritionDF$Gender == 'Female')
summary(test)

#Monthly Income is extremely left skewed, with the spike in values at or below $5000
ggplot(attritionDF, aes(MonthlyInco)) + geom_histogram(color = brewer.pal(11, "Spectral")[5], fill = brewer.pal(11, "Spectral")[4], bins = 30) + labs(title = "Monthly Income", y = "Employees", x = "USD")

#Age is slightly left skewed around 36 years of age
ggplot(attritionDF, aes(Age)) + geom_histogram(color = brewer.pal(11, "Spectral")[5], fill = brewer.pal(11, "Spectral")[4], bins = 30) + labs(title = "Employee's Ages", y = "Employees", x = "Years")
```

```{r, echo=TRUE}
#**AN** or **ANDY** or **JODI** or **TORI** please evaluate the following sentence and fill in
```

Of the employees who make less than 60,000 dollars a year, xx report low job satisfaction.


## Summary

```{r, echo=TRUE}
#**JODI** update this summary along with the ppt. Tori will also help with this.
```

Based on our research, an increase in monthly income would help you retain workers, however that is not always possible and would certainly not help the bottom line.  We would like to suggest that you explore the options of having employees set their own work schedule.  We feel that this may help eliminate some of your overtime and therefore creating workers who stay (and increase your financial status).  Are your workers working overtime because they truly need to or are they not productive during their scheduled working hours?  We would also like to suggest that you work on building capacity in your younger workers.  What are you doing to make sure they feel included in the company's day to workings.  Please consider a further review of your internal practices with younger employees.
