---
title: "EDA_Andy"
author: "Andy Ho"
date: "7/28/2018"
output: html_document
---

#### Summary: EDA on the following Variables:
+ JobRole
+ JobInvolvement
+ JobLevel
+ NumCompaniesWorked
+ PerformanceRating
+ BusinessTravel

#### Based on the following bar charts, the following stand out for high attrition:
+ Job Role: Sales Representative.  
+ Job Involvment: Low
+ Job Level: 1
+ Number of Prior Companies: Scattered so none that show as prominent.
+ Performance Rating: Both equal amount of attrition.
+ Business Travel: Those who travel frequently, which makes sense as Sales Representatives tend to travel for their role.

```{r}
library(readxl)
library(stringr)
library(ggplot2)
EDA <- read_excel("CaseStudy2-data_clean.xlsx", "HR-employee-attrition Data")
```

```{r}
# Job Role by Attrition
ggplot(EDA, aes(x = JobRole, fill = Attrition)) + geom_bar(position = "fill") + labs(title = "Job Role", x = "Job Role", y = "", color = "")+ scale_y_continuous(labels = scales::percent)+theme(axis.text.x = element_text(angle = 90))

# Job Involvement by Attrition
ggplot(EDA, aes(x = JobInvolv, fill = Attrition)) + geom_bar(position = "fill") + labs(title = "Job Involvement", x = "Job Involvement", y = "", color = "")+ scale_y_continuous(labels = scales::percent)

# Job Level by Attrition
ggplot(EDA, aes(x = JobLevel, fill = Attrition)) + geom_bar(position = "fill") + labs(title = "Job Level", x = "Job Level", y = "", color = "")+ scale_y_continuous(labels = scales::percent)

# Number of Companied Worked by Attrition
ggplot(EDA, aes(x = NumPriorComp, fill = Attrition)) + geom_bar(position = "fill") + labs(title = "Number of Prior Companies", x = "Number of Prior Companies", y = "", color = "")+ scale_y_continuous(labels = scales::percent)

# Performance Rating by Attrition
ggplot(EDA, aes(x = PerfRating, fill = Attrition)) + geom_bar(position = "fill") + labs(title = "Performance Rating", x = "Performance Rating", y = "", color = "")+ scale_y_continuous(labels = scales::percent)

#Business Travel by Attrition
ggplot(EDA, aes(x = WorkTravel, fill = Attrition)) + geom_bar(position = "fill") + labs(title = "Business Travel", x = "Business Travel", y = "", color = "")+ scale_y_continuous(labels = scales::percent)+theme(axis.text.x = element_text(angle = 90))
```
#### Creates Pie Chart of the total frequency count of Attrition
```{r}
library(plyr)
attrition.count <- count(EDA, 'Attrition')
attrition.count
slices <- attrition.count$freq
lbls <- attrition.count$Attrition
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels 
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices,labels = lbls, col=rainbow(length(lbls)),
  	main="Total Attrition for Data Set")
```

```{r}
EDA.ClassType <- as.matrix(lapply(EDA, class))
EDA.ClassType
```